Adversarial Training against Systematic Uncertainty

### 2016-09-02 tex/paper.tex
save

In this section, we show that in the setting of
Algorithm~\ref{alg:adversarial-training} where ${\cal L}_r$ is set to expected
value of the negative log-likelihood of $Z|f(X;\theta_f)$ under $r$, the
procedure converges to a classifier $f$ which is a pivotal quantity in the sense
of Eqn.~\ref{eqn:criterion}. Propositions below are derived in a non-parametric
setting, by assuming that both $f$ and $r$ have enough capacity. Results hold
for a nuisance parameter $Z$ taking either categorical values or continuous values within a
bounded support.  By abuse of notation, $H(p_Z)$ denotes the differential
entropy in this latter case. Finally, by construction, we assume the uniform
prior $p(z)$ over the nuisance values $z \in {\cal Z}$ of the parameter.

@@

\begin{proof}
For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} L_r(\theta_f,
\theta_r)$, in which case $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$ and $L_r$ reduces to the entropy
$H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
value function $E$ can therefore be rewritten as $$E'(\theta_f) = L_f(\theta_f) -
H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower bound $H(p_{Y|X}) -
H(p_Z) \leq L_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f} E'(\theta_f)$ only when
\begin{itemize}
    \item $\smash{\hat\theta}_f$ corresponds to
    the parameters of an optimal classifier, in which case the expected negative log-likelihood $L_f$ of $Y|X$
    reduces to its minimum value $H(p_{Y|X})$,
    \item all
   outcomes of $Z|f(X;\smash{\hat\theta}_f)$ are equally likely, in which case
   $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal
   Z}$ since we assume a uniform prior by construction. Note that in the continuous case, this bound is also realized
   by the uniform distribution when the support of $Z$ is bounded.
\end{itemize}
Accordingly, the second condition implies that $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$ and therefore that at this
point, because of Proposition~\ref{prop:1}, the optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal quantity.
\end{proof}

@@

>>>>

In this section, we show that in the setting of
Algorithm~\ref{alg:adversarial-training} where ${\cal L}_f$ and ${\cal L}_r$ are
respectively set to expected value of the negative log-likelihood of $Y|X$ under
$f$ and of $Z|f(X;\theta_f)$ under $r$, the procedure converges to a classifier
$f$ which is a pivotal quantity in the sense of Eqn.~\ref{eqn:criterion}.
Propositions below are derived in a non-parametric setting, by assuming that
both $f$ and $r$ have enough capacity. Results hold for a nuisance parameter $Z$
taking either categorical values or continuous values within a bounded support.
By abuse of notation, $H(p_Z)$ denotes the differential entropy in this latter
case. Finally, by construction, we assume the uniform prior $p(z)$ over the
values $z \in {\cal Z}$ of the nuisance parameter.

@@

\begin{proof}
For fixed $\theta_f$, the adversary $r$ is optimal at $\hat\theta_r = \arg
\max_{\theta_r} E(\theta_f, \theta_r)  = \arg \min_{\theta_r} L_r(\theta_f,
\theta_r)$, in which case $p_{\theta_r}(z|f(X;\theta_f)) = p(z|f(X;\theta_f))$ and $L_r$ reduces to the entropy
$H(p_{Z|f(X;\theta_f)})$ of the conditional distribution of the nuisance. The
value function $E$ can therefore be rewritten as $$E'(\theta_f) = L_f(\theta_f) -
H(p_{Z|f(X;\theta_f)}).$$  In particular, we have the lower bound $H(p_{Y|X}) -
H(p_Z) \leq L_f(\theta_f) - H(p_{Z|f(X;\theta_f)})$ where the equality
holds at $\smash{\hat\theta}_f = \arg \min_{\theta_f} E'(\theta_f)$ only when
\begin{itemize}
    \item $\smash{\hat\theta}_f$ corresponds to
    the parameters of an optimal classifier, in which case the expected negative log-likelihood $L_f$ of $Y|X$
    reduces to its minimum value $H(p_{Y|X})$,
    \item all
   outcomes of $Z|f(X;\smash{\hat\theta}_f)$ are equally likely, in which case
   $p(z|f(X;\smash{\hat\theta}_f)) = p(z)$ for all $z \in {\cal
   Z}$ since we assume a uniform prior by construction. Note that in the continuous case, the supremum of the differential entropy
   over continuous distributions of bounded support is also realized by the uniform distribution over that support.
\end{itemize}
Accordingly, the second condition implies that $p_{\theta_r}(z|f(X;\theta_f)) = p(z)$ and therefore that at this
point, because of Proposition~\ref{prop:1}, the optimal classifier $f(\cdot;\smash{\hat\theta}_f)$ is also a pivotal quantity.
\end{proof}

@@

\glnote{Define the architecture of $r$ when $Z$ is categorical ($r$ is a
standard classifier) or continuous ($r$ is e.g a 2-output NN such that the
distribution of $Z|f(X)$ is modeled by a truncated gaussian of known support).}
